---
title: "Contructing cell type-specific lncRNA-mRNA regulatory networks in ASD with Cycle"
author: "Chenchen Xiong"
date: '`r Sys.Date()`'
output:
    BiocStyle::html_document:
      toc: yes
    BiocStyle::pdf_document:
      toc: yes
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r style, echo=FALSE, results="asis", message=FALSE}
BiocStyle::markdown()
knitr::opts_chunk$set(tidy = FALSE,
    warning = FALSE,
    message = FALSE)
```

```{r echo=FALSE, results='hide', message=FALSE}
library(Cycle)
library(igraph)
```

# Introduction

As the pathogenesis of ASD involves multiple cell types and mul-tiple biological processes regulated by lncRNAs, thus it is crucial to study cell type-specific lncRNA regulation in ASD.

To explore the dynamic lncRNA regulation across different ASD cell types, we develop a novel method, Cycle, to construct cell type-specific lncRNA regulatory networks in ASD. 


# Identificaton of cell type-specific lncRNA regulation

Following the flowchart of Cycle, we could Infer cell type-specific lncRNA-mRNA regulatory networks.
Load processed matched lncRNA and mRNA snRNA-seq data with ASD samples could be downloaded from https://drive.google.com/drive/folders/1_xjA_S77POIh28w49trPnuFUCTWagBzG
```{r, eval=FALSE, include=TRUE}
load('ASD_exp_3cell_types.rda')

microglia_Cycle_networks <- Cycle_network(ASD_Microglia_ncR_data[1:100,1:100], ASD_Microglia_mR_data[1:100,1:100], boxsize = 0.1, p.value.cutoff = 0.05, num.cores = 2, dev = TRUE, iteration = TRUE, cell_id = NULL, maxiter = 20)

ASTFB_Cycle_networks <- Cycle_network(ASD_ASTFB_ncR_data[1:100,1:100],ASD_ASTFB_mR_data[1:100,1:100], boxsize = 0.1, p.value.cutoff = 0.05, num.cores = 2, dev = TRUE, iteration = TRUE, cell_id = NULL, maxiter = 20)

Neumat_Cycle_networks <- Cycle_network(ASD_Neumat_ncR_data[1:100,1:100],ASD_Neumat_mR_data[1:100,1:100], boxsize = 0.1, p.value.cutoff = 0.05, num.cores = 2, dev = TRUE, iteration = TRUE, cell_id = NULL, maxiter = 20)

celltypes = c('Microglia', 'ASTFB', 'Neumat')
Cycle_networks_tmp = list(microglia_Cycle_networks, ASTFB_Cycle_networks, Neumat_Cycle_networks)

Cycle_networks = list()
for(i in 1:length(celltypes)){
  res_list <- Cycle_networks_tmp[[i]]
  cell_num <- length(res_list)
  overlap <- Overlap.net(net = res_list, overlap.num = round(cell_num*0.9),type = 'least')
  colnames(overlap) <- c('lncRNAs','mRNAs')
  Cycle_networks[[celltypes[i]]] <- overlap
}
```


# Identifying cell type-specific hub lncRNAs 
```{r, eval=FALSE, include=TRUE}
Cycle_hubs_lncRNAs <- hub_discovery(Cycle_networks)
names(Cycle_hubs_lncRNAs) = names(Cycle_networks)
```


# Discovering conserved and rewired lncRNA regulation

Given cell type-specific lncRNA-mRNA regulatory networks, we could discover conserved and rewired lncRNA-mRNA regulatory networks.
```{r, eval=FALSE, include=TRUE}

## Overlap of multiple hubs
Overlap_net.all <- Overlap.net(Cycle_networks, overlap.num = 1, type = "least") 
Overlap_net.1 <- Overlap.net(Cycle_networks, overlap.num = 1, type = "equal")
Overlap_net.2 <- Overlap.net(Cycle_networks, overlap.num = 2, type = "least")
Overlap_net.5 <- Overlap.net(Cycle_networks, overlap.num = 5, type = "least")
Overlap_net.9 <- Overlap.net(Cycle_networks, overlap.num = 9, type = "least")
Overlap_net.15 <- Overlap.net(Cycle_networks, overlap.num = 15, type = "least")
Overlap_net.17 <- Overlap.net(Cycle_networks, overlap.num = 17, type = "equal")

Overlap_Cycle_network_rewired <- Overlap_net.1
Overlap_Cycle_network_conserved <- Overlap_net.15
```


Given cell type-specific lncRNA-mRNA regulatory networks, we could discover conserved and rewired hub lncRNAs.
```{r, eval=FALSE, include=TRUE}
## Overlap of multiple hubs
Overlap.hub.all <- Overlap.hub(Cycle_hubs_lncRNAs, overlap.num = 1, type = "least") 
Overlap.hub.1 <- Overlap.hub(Cycle_hubs_lncRNAs, overlap.num = 1, type = "equal")
Overlap.hub.2 <- Overlap.hub(Cycle_hubs_lncRNAs, overlap.num = 2, type = "least")
Overlap.hub.5 <- Overlap.hub(Cycle_hubs_lncRNAs, overlap.num = 5, type = "least")
Overlap.hub.9 <- Overlap.hub(Cycle_hubs_lncRNAs, overlap.num = 9, type = "least")
Overlap.hub.15 <- Overlap.hub(Cycle_hubs_lncRNAs, overlap.num = 15, type = "least")
Overlap.hub.17 <- Overlap.hub(Cycle_hubs_lncRNAs, overlap.num = 17, type = "equal")

Overlap_Cycle_hubs_rewired <- Overlap.hub.1
Overlap_Cycle_hubs_conserved <- Overlap.hub.15
```


# Calculating network similarity

In terms of the identified cell type-specific lncRNA-mRNA regulatory networks, we calculate network similarity of cell types.
```{r, eval=FALSE, include=TRUE}
Cycle_networks_sim <- Sim.network(Cycle_networks, Cycle_networks, directed = TRUE)
Cycle_networks_unique <- 1-Cycle_networks_sim
```


# Calculating hub similarity

In terms of the identified cell type-specific hub lncRNAs, we calculate the uniqueness of hub lncRNAs across cell types. 
```{r, eval=FALSE, include=TRUE}
Cycle_hub_Sim <- Sim.hub(Cycle_hubs_lncRNAs, Cycle_hubs_lncRNAs)
Cycle_hub_unique <- 1-Cycle_hub_Sim
```


# cell communication

communication with networks in terms of network similarity matrix between 17 cell types
```{r, eval=FALSE, include=TRUE}
Cycle_network_adjacency_matrix <- ifelse(Cycle_networks_sim > median(Cycle_networks_sim[lower.tri(Cycle_networks_sim)]), 1, 0)
diag(Cycle_network_adjacency_matrix) <- 0
colnames(Cycle_network_adjacency_matrix) <- rownames(Cycle_network_adjacency_matrix) <- names(Cycle_networks)
Cycle_network_adjacency_matrix_graph <- graph_from_adjacency_matrix(Cycle_network_adjacency_matrix, mode = "undirected")
```


Communication with hub lnRNAs in terms of hub similarity matrix between 17 cell types
```{r, eval=FALSE, include=TRUE}
## Identifying crosstalk network in terms of hub lncRNA similarity matrix between 17 cell types   
Cycle_hub_adjacency_matrix <- ifelse(Cycle_hub_Sim > median(Cycle_hub_Sim[lower.tri(Cycle_hub_Sim)]), 1, 0)
diag(Cycle_hub_adjacency_matrix) <- 0
colnames(Cycle_hub_adjacency_matrix) <- rownames(Cycle_hub_adjacency_matrix) <- names(Cycle_networks)
Cycle_hub_adjacency_matrix_graph <- graph_from_adjacency_matrix(Cycle_hub_adjacency_matrix, mode = "undirected")
```


# Session information
```{r}
sessionInfo()
```

