---
title: "Contructing cell type-specific lncRNA-mRNA regulatory networks in ASD"
author: "Chenchen Xiong"
date: '`r Sys.Date()`'
output:
    BiocStyle::html_document:
      toc: yes
    BiocStyle::pdf_document:
      toc: yes
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r style, echo=FALSE, results="asis", message=FALSE}
BiocStyle::markdown()
knitr::opts_chunk$set(tidy = FALSE,
    warning = FALSE,
    message = FALSE)
```

```{r echo=FALSE, results='hide', message=FALSE}
library(Cycle)
```


# Introduction

The rapid development of single-nucleus RNA-sequencing (snRNA-seq) technology has revolutionized the depth and resolution, which could better capture and analyze cell-specific gene expression, and deeper understand lncRNA regulation mechanism of ASD.

To explore the functions and roles of lncRNA regulation specific to different cell types in the pathogenesis of ASD, we construct cell type-specific lncRNA regulatory networks (Cycle) at the level of single cell nucleus.


# Identificaton of cell type-specific lncRNA-mRNA regulatory networks

Following the flowchart of Cycle, we could Infer cell type-specific lncRNA-mRNA regulatory networks.
```{r, eval=FALSE, include=TRUE}

load('ASD_exp_3cell_types.rda')

microglia_Cycle_network <- Cycle_network(ASD_Microglia_ncR_data, ASD_Microglia_mR_data, boxsize = 0.1, p.value.cutoff = 0.05, num.cores = 2, dev = TRUE, iteration = TRUE, cell_id = NULL, maxiter = 20)
microglia_Cycle_networks = do.call(rbind, microglia_Cycle_network)

ASTFB_Cycle_network <- Cycle_network(ASD_ASTFB_ncR_data,ASD_ASTFB_mR_data, boxsize = 0.1, p.value.cutoff = 0.05, num.cores = 2, dev = TRUE, iteration = TRUE, cell_id = NULL, maxiter = 20)
ASTFB_Cycle_networks = do.call(rbind, ASTFB_Cycle_networks)

Neumat_Cycle_network <- Cycle_network(ASD_Neumat_ncR_data,ASD_Neumat_mR_data, boxsize = 0.1, p.value.cutoff = 0.05, num.cores = 2, dev = TRUE, iteration = TRUE, cell_id = NULL, maxiter = 20)
Neumat_Cycle_networks = do.call(rbind, Neumat_Cycle_network)

```


# Discovering conserved and rewired lncRNA-mRNA regulatory networks 

Given cell type-specific lncRNA-mRNA regulatory networks, we could discover conserved and rewired lncRNA-mRNA regulatory networks.
```{r, eval=FALSE, include=TRUE}

## Overlap of multiple hubs
Overlap_net.all <- Overlap_net(Cycle_networks, overlap.num = 1, type = "least") 
Overlap_net.1 <- Overlap_net(Cycle_networks, overlap.num = 1, type = "equal")
Overlap_net.2 <- Overlap_net(Cycle_networks, overlap.num = 2, type = "least")
Overlap_net.5 <- Overlap_net(Cycle_networks, overlap.num = 5, type = "least")
Overlap_net.9 <- Overlap_net(Cycle_networks, overlap.num = 9, type = "least")
Overlap_net.17 <- Overlap_net(Cycle_networks, overlap.num = 17, type = "equal")

## Overlap of multiple hubs
Overlap.hub.all <- Overlap_hub(Cycle_hubs_lncRNAs, overlap.num = 1, type = "least") 
Overlap.hub.1 <- Overlap_hub(Cycle_hubs_lncRNAs, overlap.num = 1, type = "equal")
Overlap.hub.2 <- Overlap_hub(Cycle_hubs_lncRNAs, overlap.num = 2, type = "least")
Overlap.hub.5 <- Overlap_hub(Cycle_hubs_lncRNAs, overlap.num = 5, type = "least")
Overlap.hub.9 <- Overlap_hub(Cycle_hubs_lncRNAs, overlap.num = 9, type = "least")
Overlap.hub.17 <- Overlap_hub(Cycle_hubs_lncRNAs, overlap.num = 17, type = "equal")

Overlap_Cycle_network_rewired <- Overlap_net.1
Overlap_Cycle_hubs_rewired <- Overlap.hub.1
Overlap_Cycle_hubs_conserved <- Overlap.hub.17

```


# Calculating network similarity

In terms of the identified cell type-specific lncRNA-mRNA regulatory networks, we calculate network similarity of cell types.
```{r, eval=FALSE, include=TRUE}

library(igraph)

Cycle_networks <- list(as.data.frame(microglia_Cycle_networks), as.data.frame(ASTFB_Cycle_networks), as.data.frame(Neumat_Cycle_networks))
Cycle_networks_sim <- Sim.network(Cycle_networks, Cycle_networks, directed = TRUE)
Cycle_networks_unique <- 1-Cycle_networks_sim
```


# Calculating hub similarity

In terms of the identified cell type-specific hub lncRNAs, we calculate the uniqueness of hub lncRNAs across cell types. 
```{r}
hub1 <- list(c("AP000469.2","C2-AS1","CACNA1C-AS2"), c("DNAJC27-AS1","HPN-AS1")) 
hub2 <- list(c("CACNA1C-AS2","CALML3-AS1","CCDC148-AS1","DIAPH3-AS16"), c("LATS2-AS1", "HPN-AS1", "LINC00305"))
Cycle_hub_Sim <- Sim.hub(hub1, hub2)
Cycle_hub_unique <- 1-Cycle_hub_Sim

```


# Session information
```{r}
sessionInfo()
```

